Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-08 paper/paper.tex
Move C- section Change DFD diagram to be left -> right

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}

@@

For example, the programs in Fig.~\ref{fig:c-} are valid
\newC programs, whereas the program in Fig.~\ref{fig:not-c-} is not, since it uses
a non-constant loop bound.

\begin{figure}
 \begin{lstlisting}[language=C]
int fib(int n) {
  int a, b, c;

  a = b = 1;

  while (n--) {
    c = a+b;
    a = b;
    b = c;
  }

  return a;
}
 \end{lstlisting}

 \caption{A C program to compute Fibonacci numbers.}
  \label{fig:not-c-}

\subsection{Why use \newC instead of SAT?}
To misquote David Wheeler: ``all problems in computer science can be solved by
another level of abstraction.''~\cite{beautiful-code}.  \newC serves as
an abstraction of SAT.  We can think of SAT as an assembly language and
{\sc cbmc} as a compiler for \newC to this assembly language.  The benefits
of using \newC as our logic rather than SAT are very similar to the benefits of
writing code in a high level language rather than assembly language.
Some of the more significant advantages are:

@@

\subsection{The Concrete Algorithm for Bitvector Programs}

@@

In order to determine the validity of the {\sc synth} formula, we can now
check the {\sc synth} program for safety.  The {\sc synth} program is a
\newC program, which means we can check its safety by invoking a bounded
model checker, such as {\sc cbmc}.  Since \newC is a fragment of C,
the {\sc synth} program can also be compiled using {\sc gcc}.
We can add a small piece of code which enumerates each possible
$\mathcal{L}$-program, then evaluates each program on the current test vectors by
running {\sc synth}. If it finds a program which is correct, it emits
the program as a candidate, just like a conventional superoptimiser.
The combination of this enumeration function and {\sc synth} can be compiled
to an executable using {\sc gcc}.  This leaves
us with a very simple explicit-state model checker which decides
safety for {\sc synth}.  We can run the explicit-state model checker
in parallel with {\sc cbmc} and take the answer of whichever happens
to terminate first, stopping the other process.  This procedure is
depicted for {\sc synth} in Fig.~\ref{fig:synth-dfd} and for
{\sc verif} in Fig.~\ref{fig:verif-dfd}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[
  scale=0.8,
  font=\sffamily,
  every matrix/.style={ampersand replacement=\&,column sep=.5cm,row sep=.75cm},
  source/.style={draw,thick,rounded corners,fill=yellow!20,inner sep=.3cm},
  process/.style={draw,thick,circle,fill=blue!20},
  sink/.style={source,fill=green!20},
  datastore/.style={draw,very thick,shape=datastore,inner sep=.3cm},
  dots/.style={gray,scale=2},
  to/.style={->,>=stealth',shorten >=1pt,semithick,font=\sffamily\footnotesize},
  every node/.style={align=center, scale=0.8}]

  % Position the nodes using a matrix layout
  \matrix{
    \& \node[source] (driver) {driver}; \& \node[source] (user) {user}; \& \\
    \& \node[datastore] (tests) {\texttt{tests.c}}; \& \node[datastore] (spec) {\texttt{spec.c}}; \& \node[datastore] (interpreter) {\texttt{interpreter.c}}; \\
    \node[datastore] (cbmc-synth) {\texttt{cbmc/synth.c}}; \& \node[process] (cbmc) {\sc cbmc}; \& \node[process] (gcc) {\sc gcc}; \& \node[datastore] (explicit-synth) {\texttt{explicit/synth.c}}; \\
    \& \& \node[process] (aout) {\texttt{a.out}}; \& \\
    \& \node[sink] (cbmc-out) {candidate program}; \& \node[sink] (explicit-out) {candidate program}; \& \\
  };

  \draw[to] (driver) -- (tests);
  \draw[to] (user) -- (spec);

  \draw[to] (tests) -- (cbmc);
  \draw[to] (spec) -- (cbmc);
  \draw[to] (interpreter) -- (cbmc);

  \draw[to] (tests) -- (gcc);
  \draw[to] (spec) -- (gcc);
  \draw[to] (interpreter) -- (gcc);

  \draw[to] (cbmc-synth) -- (cbmc);
  \draw[to] (explicit-synth) -- (gcc);

  \draw[to] (gcc) -- (aout);

  \draw[to] (cbmc) -- (cbmc-out);
  \draw[to] (aout) -- (explicit-out);
\end{tikzpicture}
\end{center}

@@

\begin{figure}
\begin{center}
\begin{tikzpicture}[
  scale=0.8,
  font=\sffamily,
  every matrix/.style={ampersand replacement=\&,column sep=.5cm,row sep=.75cm},
  source/.style={draw,thick,rounded corners,fill=yellow!20,inner sep=.3cm},
  process/.style={draw,thick,circle,fill=blue!20},
  sink/.style={source,fill=green!20},
  datastore/.style={draw,very thick,shape=datastore,inner sep=.3cm},
  dots/.style={gray,scale=2},
  to/.style={->,>=stealth',shorten >=1pt,semithick,font=\sffamily\footnotesize},
  every node/.style={align=center, scale=0.8}]

  % Position the nodes using a matrix layout
  \matrix{
    \& \node[source] (driver) {driver}; \& \node[source] (user) {user}; \& \\
    \& \node[datastore] (prog) {\texttt{prog.c}}; \& \node[datastore] (spec) {\texttt{spec.c}}; \& \node[datastore] (interpreter) {\texttt{interpreter.c}}; \\
    \node[datastore] (cbmc-verif) {\texttt{cbmc/verif.c}}; \& \node[process] (cbmc) {\sc cbmc}; \& \node[process] (gcc) {\sc gcc}; \& \node[datastore] (explicit-verif) {\texttt{explicit/verif.c}}; \\
    \& \& \node[process] (aout) {\texttt{a.out}}; \& \\
    \& \node[sink] (cbmc-out) {test vector}; \& \node[sink] (explicit-out) {test vector}; \& \\
  };

  \draw[to] (driver) -- (prog);
  \draw[to] (user) -- (spec);

  \draw[to] (prog) -- (cbmc);
  \draw[to] (spec) -- (cbmc);
  \draw[to] (interpreter) -- (cbmc);

  \draw[to] (tests) -- (gcc);
  \draw[to] (spec) -- (gcc);
  \draw[to] (interpreter) -- (gcc);

  \draw[to] (cbmc-verif) -- (cbmc);
  \draw[to] (explicit-verif) -- (gcc);

  \draw[to] (gcc) -- (aout);

  \draw[to] (cbmc) -- (cbmc-out);
  \draw[to] (aout) -- (explicit-out);
\end{tikzpicture}
\end{center}

\caption{Schematic diagram of {\sc verif}}
\label{fig:verif-dfd}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=\linewidth]{figures/brahma_kalashnikov.pdf}
 \caption{Brahma runtimes vs. Kalashnikov}
  \label{fig:brahma-kalashnikov}

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
%\usepackage{algorithmic}

@@

@@

\subsection{Why use \newC instead of SAT?}
To misquote David Wheeler: ``all problems in computer science can be solved by
another level of abstraction''~\cite{beautiful-code}.  \newC serves as
an abstraction of SAT.  We can think of SAT as an assembly language and
{\sc cbmc} as a compiler for \newC to this assembly language.  The benefits
of using \newC as our logic rather than SAT are very similar to the benefits of
writing code in a high level language rather than assembly language.
Some of the more significant advantages are:

\section{The Concrete Algorithm for Bitvector Programs}

@@

In order to determine the validity of the {\sc synth} formula, we can now
check the {\sc synth} program for safety.  The {\sc synth} program is a
\newC program, which means we can check its safety by invoking a bounded
model checker, such as {\sc cbmc}.  Since \newC is a fragment of C,
the {\sc synth} program can also be compiled using {\sc gcc}.
We can add a small piece of code which enumerates each possible
$\mathcal{L}$-program, then evaluates each program on the current test vectors by
running {\sc synth}. If it finds a program which is correct, it emits
the program as a candidate, just like a conventional superoptimiser.
The combination of this enumeration function and {\sc synth} can be compiled
to an executable using {\sc gcc}.  This leaves
us with a very simple explicit-state model checker which decides
safety for {\sc synth}.  We can run the explicit-state model checker
in parallel with {\sc cbmc} and take the answer of whichever happens
to terminate first, stopping the other process.  This procedure is
depicted for {\sc synth} in Fig.~\ref{fig:synth-dfd}, and is similar for {\sc verif}.

\begin{figure}
\begin{center}
\tikzstyle{file}=[draw, text width=7.0em, text centered,
  minimum height=1.5em]
\tikzstyle{process} = [draw, minimum height=3em, circle]
\tikzstyle{line} = [draw, color=black, -latex']

\begin{tikzpicture}[font=\sffamily]

\node [file] (synth) {\sc synth.c};
\path (synth.south)+(0.0, -0.5) node [file] (tests) {\sc tests.c};
\path (tests.south)+(0.0, -0.5) node [file] (interpreter) {\sc interpreter.c};
\path (interpreter.south)+(0.0, -0.5) node [file] (spec) {\sc spec.c};

\path (tests.east)+(2.0, -0.25) node [process] (merged) {merge};

\path (merged.east)+(2.0, 1.0) node [process] (cbmc) {\sc cbmc};
\path (merged.east)+(2.0, -1.0) node [process] (gcc) {\sc gcc};

\path (cbmc.east)+(2.5, -1.0) node [file] (out) {candidate program};

\path [line] (synth) -- (merged);
\path [line] (tests) -- (merged);
\path [line] (interpreter) -- (merged);
\path [line] (spec) -- (merged);

\path [line] (merged) -- (cbmc);
\path [line] (merged) -- (gcc);

\path [line] (cbmc) -- (out);
\path [line] (gcc) -- (out);

\end{tikzpicture}
\end{center}

@@

@@

%\begin{figure}
% \includegraphics[width=\linewidth]{figures/brahma_kalashnikov.pdf}
% \caption{Brahma runtimes vs. Kalashnikov}
%  \label{fig:brahma-kalashnikov}

%\end{figure}
